---
title: "Data cleaning"
author: "Payalben Siddharthsinh chavda"
date: "June 9, 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(ggplot2)
library(plyr)
```

## Data cleaning 

## Read data file into R
```{r}
df <- read.csv("C:/Users/pchav/Downloads/MCI_2014_to_2019.csv")
```
## Drop irrelevant attributes
```{r}
crime_data <- df[,c(5,15:21,24:26)]
```
## Let’s check the structure of dataset:

```{r}
str(crime_data)
print("All variables have correct data type. There are 5 categorical and 6 numerical attributes.")
```


## summary of dataset 
```{r}
summary(crime_data)
print("summary shows that there is missing data in occurrenceyear, occurrencedayofyear, occurrenceday has missing values.
")
```

## analyze categorical variables by frequency table
```{r}
count(crime_data, "occurrencemonth")
count(crime_data,"occurrencedayofweek")
count(crime_data,"MCI")
count(crime_data,"Neighbourhood")
count(crime_data,"premisetype")

print("Frequency tabble shows that occurrencemonth and occurrencedayofweek has missing values.")
```
 
 
## check records with missing values
```{r}
missing_records <- crime_data[is.na(crime_data$occurrenceyear),]
print("AS we look at those 59 records it contains 5 variable with missing values. So it’s better to remove those records as they won’t provide any useful information.")
```


## Dealing with missing values
```{r}
crime_data <- crime_data[!is.na(crime_data$occurrenceyear),]
sum(is.na(crime_data))
```

## Outlier detection using boxplot:

```{r}
boxplot(crime_data$occurrenceyear)
boxplot(crime_data$occurrenceday)
boxplot(crime_data$occurrencehour)
boxplot(crime_data$occurrencedayofyear)
boxplot(crime_data$Long)
boxplot(crime_data$Lat)
print("This shows that only “occurrenceyear” has outliers.As project aim to analyse 2014-2019 data, we will remove records with of 2000 to 2013.")
```


##Removing records from year 2000 to 2013:

```{r}
crime_data <- crime_data[(crime_data$occurrenceyear > 2013),]
```

## Remove duplicates from dataset
```{r}
crime_data <- crime_data[!duplicated(crime_data), ]

print("This is our consistence dataset.I will use this dataset for futher analysis.")
```



## Dimension of consistence dataset
```{r}
dim(crime_data)
```

## Imbalance data
```{r}
prop.table(table(crime_data$MCI))

print("This shows data is imbalance, more than 50 percent of data belongs to Assault class.")
```

